Language-Based Security for Low-Level MPC

### 2024-05-17 ppdp24/intro.tex
checkpoint

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
but mostly manual. Somewhat independently, a significant
body of work in programming languages has focused on definition and
enforcement of confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and delimited
release \cite{XXX}. But following a tradition of connecting
cryptographic and PL-based security models \cite{XXX}, recent work has
also recognized connections between MPC security models and
hyperpoperties of, e.g., noninterference \cite{XXX}, and even
leveraged these connections to enforce MPC security through mechanisms
such as security types \cite{XXX}. Major benefits of this connection
in an MPC setting include better language abstractions for defining
protocols and for mechanisation and even automation of security proofs.

The goal of this paper is to develop a PL model for defining a variety of
low-level probabilistic MPC protocols, to formulate a collection
of confidentiality and integrity hyperproperties for our model
model with familiar information flow analogs, and to show how these
properties can be leveraged for improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria \cite{XXX} and
Viaduct \cite{XXX} are designed to provide effective programming of
full applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration \cite{XXX} to
guarantee high-level security properties, and they rely on
\emph{libraries} of low-level MPC protocols, such as binary and
arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and rely on probabilistic progamming methods. So, low-level MPC
programming and protocol verification is a distinct challenge but
critical to the general challenge of PL design for MPC, and complementary
to high-level language design.

Analysis of

have been developed

The security
threat models and engMPC is mainly the
product of a significant body of work in cryptography,

information flow security has received
significant attention in PL theory and practice especially since the
turn of the century \cite{1159651}, including a menagerie of variants,
enforcement mechanisms, and programming frameworks. Much of this has
been enabled by the unified metatheory of \emph{hyperproperties}
\cite{10.5555/1891823.1891830}, including noninterference, that
establishes a common conceptual framework for reasoning about and
implementing systems with information flow security.

On its surface, security for MPC protocols seems similar to
probabilistic noninterference. However, MPC security differs from
noninterference in a subtle but fundamental way, and approaches for
verifying noninterference cannot naturally extend to MPC security. Our
goal is to explore connections between the security model of MPC-- the
\emph{real/ideal} aka \emph{simulator security} model -- and
trace-based hyperproperties, and to leverage these connections to
obtain automated enforcement mechanisms for MPC protocol
development. Currently, proof methods for MPC protocol development are
well-studied \cite{Lindell2017} with some previous work leveraging
connections with hyperproperties \cite{8429300}. But these methods
require fully manual proof construction and are thus tedious and
error-prone. Therefore our exploration makes both theoretical and
practical contributions.

\subsection{The Analysis Challenge of MPC}
MPC protocols involve communication
between a group of distributed participants called a \emph{federation}
that collaboratively compute and publish the result of some known
\emph{ideal functionality} $\idealf$, while keeping each party's input
``secret'', without the use of a trusted third party. This last part
is critical. For example, if we take $\idealf$ to be the majority vote
function, a protocol for computing $\idealf$ is MPC-secure if, given
any set of input votes, it correctly computes and publishes the voting
result but reveals no other information to the public or to other
participants. However, by publishing the result, some information
about individual votes may be implicitly declassified.  For example,
in the case of a majority vote in a federation of size 3, if the
motion carries and party 1 has voted no, then party 1 knows exactly
the votes of parties 2 and 3. This cannot be avoided due to the nature
of $\idealf$.

Security in the MPC setting thus means that protocols cannot reveal
any secret information other than what is implicitly declassified by
the publicized output of the ideal functionality. The security model
also assumes that some subset of participants can be corrupted and
collude adversarially to possibly infer more secret information. The
accepted method of demonstrating protocol security in this setting is
to define a \emph{simulation} algorithm that runs in the ``ideal''
world which, given just the inputs of corrupted parties and the output
of the ideal functionality, is able to reconstruct information that
corrupted parties receive in their so-called \emph{views} of the
protocol running in the real world.  This implies that adversarial
views provide no information beyond what is provided by the ideal
output alone. Simulation is defined probabilistically since MPC
protocols typically rely on cryptographic and probabilistic methods.

In Section \ref{section-hyperprop-passive} we formalize real/ideal
security, and we define and discuss simple examples of MPC protocols
in Section \ref{section-minicat-examples} and more complex ones in
Section \ref{section-metalang}. But an immediate and main point in
relation to security hyperproperties is that, due to the potential for
information release in MPC, simulator security is \emph{not} a strict
probabilistic noninterference or trace obliviousness property, as we
show in Section \ref{section-hyperprop-ni}- rather, the public output
allows and sets an upper bound on declassification.

@@

Our main focus is on automated and semi-automated reasoning about
security properties of specific protocols, both through whole-program
analysis and compositional properties of program components that
support security.

Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. This work is explicitly
reminiscent of information flow approaches such as delimited
information release \cite{10.1007/978-3-540-37621-7_9}, downgrading
policies \cite{li2005downgrading}, and relaxed noninterference
\cite{10.1145/1040305.1040319}. However, their program logic assumes
correctness of the underlying MPC protocols.  The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. However their language ultimately compiles to
protocols assumed to be secure.  All of these systems leverage type
systems to obtain automated security analyses.

In the realm of protocol verification, previous work has shown how to
reduce proofs of simulator security to program equivalence assertions,
which supports computer-aided verification of MPC protocols in
EasyCrypt \cite{8429300}. This work also leveraged hyperproperty
characterizations of simulator security, both passive and
active. However, their EasyCrpyt security proof constructions are
entirely manual. Probabilistic Separation Logic
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence in programs.  They consider
several MPC protocols and show how to systematically deduce, e.g.,
probabilistic independence of input secrets and adversarial views that
are critical security properties.  However, their methods are manual. 
Also, their focus on independence does not address the subtleties of
dependence that arise during MPC protocol execution, and the authors
do not formulate or prove simulator security properties in
\cite{barthe2019probabilistic}. The $\lambda_{\mathrm{obliv}}$
language \cite{darais2019language} is perhaps most closely related to
our work. Here, a type system is used to automatically enforce
so-called probabilistic trace obliviousness and they show how to use
their system to implement several variants of oblivious RAM. However,
obliviousness is related to noninterference and does not address
allowable information leakage as in MPC, and their use of type
linearity disallows the reuse of sampled probabilistic values which is
common in the MPC idiom, as our examples will illustrate. Similarly,
previous work on oblivious data structures \cite{10.1145/3498713},
while intended for sound integration into MPC protocols, focus on
expression and enforcement of obliviousness and do not account for
allowable information leakage.

Our work also shares many ideas with probabilistic programming
languages designed to perform (exact or approximate)
inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
verifying properties beyond inference, including conditional
independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
manual approach for proving such properties, but does not provide
automation.

>>>>

Secure Multi-Party Computation (MPC) protocols support data privacy in
important modern, distributed applications such as privacy-preserving
machine learning \cite{li2021privacy, knott2021crypten,
  koch2020privacy, liu2020privacy} and Zero-Knowledge proofs in
blockchains \cite{ishai2009zero, lu2019honeybadgermpc,
  gao2022symmeproof, tomaz2020preserving}. The security semantics of
MPC include both confidentiality and integrity properties incorporated
into models such as real/ideal (aka simulator) security and universal
composability (UC), developed primarily by the cryptography community
\cite{XXX}.  Related proof methods are well-studied \cite{Lindell2017}
but mostly manual. Somewhat independently, a significant
body of work in programming languages has focused on definition and
enforcement of confidentiality and integrity \emph{hyperproperties}
\cite{10.5555/1891823.1891830} such as noninterference and delimited
release \cite{XXX}. Following a tradition of connecting
cryptographic and PL-based security models \cite{XXX}, recent work has
also recognized connections between MPC security models and
hyperpoperties of, e.g., noninterference \cite{8429300}, and even
leveraged these connections to enforce MPC security through mechanisms
such as security types \cite{XXX}. Major benefits of this connection
in an MPC setting include better language abstractions for defining
protocols and for mechanisation and even automation of security proofs.
The goal of this paper is to develop a PL model for defining a variety of
low-level probabilistic MPC protocols, to formulate a collection
of confidentiality and integrity hyperproperties for our model
model with familiar information flow analogs, and to show how these
properties can be leveraged for improved proof automation.

The distinction between high- and low-level languages for MPC is
important. High-level languages such as Wysteria \cite{XXX} and
Viaduct \cite{XXX} are designed to provide effective programming of
full applications. These language designs incorporate sophisticated
verified compilation techniques such as orchestration \cite{XXX} to
guarantee high-level security properties, and they rely on
\emph{libraries} of low-level MPC protocols, such as binary and
arithmetic circuits. These low-level protocols encapsulate
abstractions such as secret sharing and semi-homomorphic encrypytion,
and rely on probabilistic progamming methods. So, low-level MPC
programming and protocol verification is a distinct challenge and both
critical to the general challenge of PL design for MPC and
complementary to high-level language design.

The connection between information flow hyperproperties and
MPC security is also complicated especially at a low level.
MPC protocols involve communication between a group of distributed
clients called a \emph{federation} that collaboratively compute
and publish the result of some known \emph{ideal functionality}
$\idealf$, maintaining confidentiality of inputs to $\idealf$ without
the use of a trusted third party. However, since the outputs of
$\idealf$ are public, some information about inputs is inevitably
leaked. Thus, the ideal
functionality establishes a declassification policy \cite{XXX}, which
is more difficult to enforce than pure noninterference.  And
subtleties of, e.g., semi-homomorphic encryption are central to both
confidentiality and integrity properties of protocols and similarly
difficult to track with coarse-grained security types alone.

Nevertheless, as previous authors have observed \cite{XXX}, low-level
protocol design patterns such as secret sharing and circuit gate
structure have compositional properties that can be abstracted and
leveraged in larger proof contexts. We contribute to this line of work
by developing an automated verification technique for subprotocols and
show how it can be integrated as a tactic in a larger security proof.

Our main focus is on PL design and automated and semi-automated
reasoning about security properties of low-level MPC protocols.  We
consider related systems in several dimensions, including
whether they are aimed at low-level design with probabilistic features,
whether they support reasoning about conditional probabilities which
are central to real/ideal security, whether they consider MPC
security through the lens of hyperproperties, and whether they consider
passive and/or malicious security models. We summarize this comparison
in Figure \ref{fig-comp}, with the caveat that works vary in the degree
of development in each dimension.

As mentioned above, several high-level languages have been developed
for writing MPC applications, 
and frequently exploit the connection between hyperproperties and MPC
security. Previous work on analysis for the SecreC language
\cite{almeida2018enforcing,10.1145/2637113.2637119} is concerned with
properties of complex MPC circuits, in particular a user-friendly
specification and automated enforcement of declassification bounds in
programs that use MPC in subprograms. The Wys$^\star$ language
\cite{wysstar}, based on Wysteria \cite{rastogi2014wysteria}, has
similar goals and includes a trace-based semantics for reasoning about
the interactions of MPC protocols. Their compiler also guarantees that
underlying multi-threaded protocols enforce the single-threaded source
language semantics. These two lines of work were focused on passive
security. The Viaduct language \cite{XXX} has a well-developed
information flow type system that automatically enforces both
confidentiality and integrity through hyperproperties such as robust
declassification, in addition to rigorous compilation guarantees
through orchestration \cite{XXX}. However, these high level
languages lack probabilistic features and other abstractions of
low-level protocols, the implementation and security of which
are typically assumed as a selection of library compnonents.

Various related low-level languages with probabilistic features have
also been developed. The $\lambda_{\mathrm{obliv}}$ language
\cite{darais2019language} uses a type system is used to automatically
enforce so-called probabilistic trace obliviousness.  But similar to
previous work on oblivious data structures \cite{10.1145/3498713},
obliviousness is related to pure noninterference, not the relaxed form
related to passive MPC security. The Haskell-based security type system in
\cite{XXX} enforces a version of conditioned noninterence that is
sound for passive security in a probabilisitic low-level setting, but
they do not consider malicious security. And properties of real/ideal
passive and malicious security for a probabilistic language have been
formulated in EasyCrypt \cite{8429300}-- though their proof methods,
while mechanized, are fully manual, and their formulation of malicious
security is not as clearly related to robust declassification as the
one we present in Section \ref{section-hyperprop}.

Program logics for probabilistic languages and specifically reasoning
about properties such as joint probabilistic independence is also
important related work. Probabilistic Separation Logic (PSL)
\cite{barthe2019probabilistic} develops a logical framework for
reasoning about probabilistic independence (aka separation) in
programs, and they consider several (hyper)properties, such as perfect
secrecy of one-time-pads and indistinguishability in secret sharing,
that are critical to MPC. However, their methods are manual, and they
do not capture notions of conditional independence (separation). This
latter issue has been addressed in Lilac \cite{XXX}, but the
application of Lilac-style reasoning to MPC protocols has not
previously been explored, as we do in Section
\ref{section-example-gmw}.

Our work also shares many ideas with probabilistic programming
languages designed to perform (exact or approximate) statistical
inference~\cite{holtzen2020scaling, carpenter2017stan, wood2014new,
  bingham2019pyro, albarghouthi2017fairsquare, de2007problog,
  pfeffer2009figaro, saad2021sppl}. Our setting, however, requires
verifying properties beyond inference, including conditional
statistical independence. Recent work by Li et al.~\cite{li2023lilac} proposes a
manual approach for proving such properties, but does not provide
automation.

@@
